API design for machine learning software: experiences from the scikit-learn project

### 2013-06-27 paper.tex
Make subsection for the general principles.
Should hopefully address @amueller's concern.
Improving the tone in the comparisons section and the future work section

All objects within \sklearn share a uniform common basic API consisting of three
complementary interfaces: an \textit{estimator} interface for building and
fitting models, a \textit{predictor} interface for making predictions and a
\textit{transformer} interface for converting data. As much as possible, our
design choices have been guided so as to avoid the proliferation of framework
code. We try to adopt simple conventions and to limit to a minimum the number of
methods an object must implement. The API is designed to adhere to the following
broad principles:

@@

In \sklearn, classical learning algorithms are not the only objects to be
implemented as estimators. For example, preprocessing routines (e.g., scaling of
features) or feature extraction techniques (e.g., vectorization of text
documents) also implement the \textit{estimator} interface. Even processing steps that are
entirely stateless and therefore do not require the \texttt{fit} method to
perform any useful work implement the estimator interface. As we will illustrate
in the next sections, this design pattern is indeed of prime importance for
consistency, composition and model selection reasons.

@@

Recent years have witnessed a rising interest in machine learning and data mining
with applications in many fields.
With this rise comes a host of machine learning packages
(both open source and proprietary) with which \sklearn competes.
Although it can also be used through its Java API, Weka \citep{hall2009weka} is
popularly used through its graphical user interface (GUI). We argue that a GUI
lacks flexibility, makes experiments difficult to reproduce and has a high
maintenance cost from a development point of view. For these reasons, we make no
attempt to provide such a GUI in \sklearn.
Other existing machine learning packages 
such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
and Vowpal~Wabbit\footnote{\url{http://hunch.net/$\sim$vw/}}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process input/output,
and will do so in a variety of languages.
We argue that this makes it harder for researchers and developers
to exchange and collaborate on software.

In comparison to specialized languages for numeric and statistical programming
such as \textsc{matlab} and R \citep{r},
Python has the distinct advantage
that it is a \textit{general purpose} language,
while NumPy and SciPy extend it with functionality
similar to that offered by \textsc{matlab} and R.
Python has strong language and standard library support for such tasks as
string/text processing, interprocess communication, networking
and many of the other tasks that machine learning programs
(whether academic or commercial) routinely need to perform.
However, \textsc{matlab} and R remain the preferred language of many
researchers. As such, much scientific software is still open-sourced primarily
in these two languages.

@@

\section{Future work}
\label{sec:future_work}
There are several problems with the current \sklearn interface which
will be resolved in the future.
At present, the library does not support some classical machine learning
algorithms,
including neural networks, ensemble meta-estimators for
bagging or subsampling strategies and missing value completion algorithms.
Tasks like structured prediction or reinforcement learning are
considered off-scope for the project, since they would require quite different
data representations and APIs.

>>>>

All objects within \sklearn share a uniform common basic API consisting of three
complementary interfaces: an \textit{estimator} interface for building and
fitting models, a \textit{predictor} interface for making predictions and a
\textit{transformer} interface for converting data. In this section, we describe
these three interfaces, after reviewing our general principles and data
representation choices.

\subsection{General principles}

As much as possible, our
design choices have been guided so as to avoid the proliferation of framework
code. We try to adopt simple conventions and to limit to a minimum the number of
methods an object must implement. The API is designed to adhere to the following
broad principles:

@@

In \sklearn, classical learning algorithms are not the only objects to be
implemented as estimators. For example, preprocessing routines (e.g., scaling of
features) or feature extraction techniques (e.g., vectorization of text
documents) also implement the \textit{estimator} interface. Even stateless
processing steps, that do not require the \texttt{fit} method to
perform useful work, implement the estimator interface. As we will illustrate
in the next sections, this design pattern is indeed of prime importance for
consistency, composition and model selection reasons.

@@

Recent years have witnessed a rising interest in machine learning and data mining
with applications in many fields.
With this rise comes a host of machine learning packages
(both open source and proprietary) with which \sklearn competes.
Although it can also be used through its Java API, Weka \citep{hall2009weka} is
popularly used through its graphical user interface (GUI). The target audience
of the \sklearn is capable of programming, and therefore we focus on developing
a usable and consistent API, rather than expend effort into creating a GUI. In
addition, running machine learning through a GUI makes reproducibility difficult
(although Weka has developed a GUI for managing such workflows).

Other existing machine learning packages 
such as SofiaML\footnote{\url{http://code.google.com/p/sofia-ml/}}
and Vowpal~Wabbit\footnote{\url{http://hunch.net/$\sim$vw/}}
are intended to be used as command-line tools
(and sometimes do not offer any type of API).
While these packages have the advantage
that their users are not tied to a particular programming language,
the users will find that they still need programming to process input/output,
and will do so in a variety of languages.
The focus of \sklearn to incorporate the entire package into a single package still
allows for the use of external tools to process input/output, however gives the added
benefit of being able to run these processes within the same development environment.
In addition, this makes it easier for researchers and developers
to exchange and collaborate on software, as dependencies and setup are kept to a
minimum.

In comparison to specialized languages for numeric and statistical programming
such as \textsc{matlab} and R \citep{r},
Python has the distinct advantage
that it is a \textit{general purpose} language,
while NumPy and SciPy extend it with functionality
similar to that offered by \textsc{matlab} and R.
Python has strong language and standard library support for such tasks as
string/text processing, interprocess communication, networking
and many of the other auxiliary tasks that machine learning programs
(whether academic or commercial) routinely need to perform.
While support for many of these tasks is improving in languages such as
\textsc{matlab} and R, they still lag behind Python in their general purpose
applicability.
In many applications of machine learning these tasks, such as data access,
data preprocessing and reporting, can be a more significant task than applying 
the actual learning algorithm.

@@

\section{Future directions}
\label{sec:future_work}
There are several directions that the \sklearn project
aims to focus on in future development.
At present, the library does not support some classical machine learning
algorithms,
including neural networks, ensemble meta-estimators for
bagging or subsampling strategies and missing value completion algorithms.
Tasks like structured prediction or reinforcement learning are
considered off-scope for the project, since they would require quite different
data representations and APIs.
